---
title: "IS607 WK6 Assignment"
author: "J. Hamski"
date: "March 7, 2015"
output: html_document
---
***
1. Choose and load any R dataset (except for diamonds!) that has at least two numeric variables and at least two categorical variables.  Identify which variables in your data set are numeric, and which are categorical (factors).
2. Generate summary level descriptive statistics:  Show the mean, median, 25th and 75th quartiles, min, and max for each of the applicable variables in your data set.
3. Determine the frequency for one of the categorical variables.
4. Determine the frequency for one of the categorical variables, by a different categorical variable.
5. Create a graph for a single numeric variable.
6. Create a scatterplot of two numeric variables.

> Packages used: datasets, ggplot2, dplyr, tidyr

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(datasets)
library(ggplot2)
library(dplyr)
library(tidyr)
```
```{r}
data(mtcars)

str(mtcars)
```

While all data are of the class numeric, number of cynlinders (cyl), ? (vs), transmission type (am), number of forward gears (gear) and number of carburetor (carb) are better described as categorical variables.
